04_DESeq_sexes_manipulate

### 2019-12-02 analysis/04_DESeq2_sexes_manipulate.Rmd
now with manip volcanos

# rename some factors
levels(c.colData$treatment)
# "extend"    "m.inc.d17" "m.inc.d3"  "m.inc.d8"  "m.inc.d9"  "m.n2"      "prolong" 
c.colData <- c.colData %>%
    mutate(treatment = fct_recode(treatment,
                            "extend" = "extend",
                            "prolong" = "prolong",
                            "early" = "m.inc.d8",
                            "remove.d03" = "m.inc.d3",
                            "remove.d09" = "m.inc.d9",
                            "remove.d17" = "m.inc.d17",
                            "remove.d20" = "m.n2"))

c.colData$treatment <- factor(c.colData$treatment, levels = maniplevels)
levels(c.colData$treatment)

# drop columsn that are bad for model AND not needed for later joining
c.colData$bird <- NULL
c.colData$tissue <- NULL
c.colData$study <- NULL
c.colData$group <- NULL
head(c.colData)
str(c.colData)
```

@@

>>>>

# rename some factors
levels(c.colData$treatment)
# "extend"    "m.inc.d17" "m.inc.d3"  "m.inc.d8"  "m.inc.d9"  "m.n2"      "prolong"

c.colData$treatment <- factor(c.colData$treatment, levels = maniplevels1)
levels(c.colData$treatment)

# drop columsn that are bad for model AND not needed for later joining
c.colData$bird <- NULL
c.colData$tissue <- NULL
c.colData$study <- NULL
c.colData$group <- NULL
head(c.colData)
str(c.colData)

geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)
```

@@

## volcano plots

```{r pitdfs, eval=T}
dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary","male_pituitary"))
dds.pituitary$treatment
```

```{r pitdfs2, eval=T}
createDEGdfmanip <- function(mydds, whichfactor, up, down, mytissue){
  res <- results(mydds, contrast =c(whichfactor, up, down),
                 independentFiltering = T, alpha = 0.1)
   data <- data.frame(entrezid = row.names(res),
                     padj = res$padj, 
                     logpadj = -log10(res$padj),
                     lfc = res$log2FoldChange,
                     tissue = mytissue)
  data <- na.omit(data)

  data <- data %>%
    dplyr::mutate(direction = ifelse(data$lfc > 0 & data$padj < 0.1, 
                                     yes = up, no = ifelse(data$lfc < 0 & data$padj < 0.1, 
                                                 yes = down, no = "NS")))
  data$direction <- factor(data$direction, levels = c(down, "NS", up))

  data <- left_join(data, geneinfo) %>%
    dplyr::mutate(gene = Name) %>%
    dplyr::select(gene, lfc, padj, logpadj, direction, tissue) %>%
    dplyr::arrange(desc(lfc))

  DEGs <- data %>% filter(direction != "NS")
  print(str(DEGs))

  myfilename = paste("../results/04_DEGs", mytissue, down,up, "csv", sep = ".")

  # write dataframe of only significant genes
  write.csv(DEGs, myfilename, row.names = F)

  # return data frome with all data, included NS genes
  return(data)
}

m.inc.d3.m.inc.d9 <- createDEGdfmanip(dds.pituitary, "treatment", "m.inc.d9", "m.inc.d3",  "pituitary")
m.inc.d9.m.inc.d17 <- createDEGdfmanip(dds.pituitary, "treatment", "m.inc.d17",  "m.inc.d9", "pituitary")
m.inc.d17.m.n2 <- createDEGdfmanip(dds.pituitary, "treatment", "m.n2",  "m.inc.d17", "pituitary") 
m.n2.m.inc.d8 <- createDEGdfmanip(dds.pituitary, "treatment", "m.inc.d8",  "m.n2", "pituitary") 
m.inc.d8.prolong <- createDEGdfmanip(dds.pituitary,  "treatment", "prolong", "m.inc.d8",  "pituitary") 
prolong.extend <- createDEGdfmanip(dds.pituitary, "treatment", "extend", "prolong",  "pituitary")
```

```{r pitplot, eval = T}
a <- plot.volcano(m.inc.d3.m.inc.d9, "treatment", "m.inc.d9", "m.inc.d3",  colorsvolcano)
b <- plot.volcano(m.inc.d9.m.inc.d17, "treatment", "m.inc.d17", "m.inc.d9",  colorsvolcano)
c <- plot.volcano(m.inc.d17.m.n2, "treatment",  "m.n2", "m.inc.d17", colorsvolcano) 
d <- plot.volcano(m.n2.m.inc.d8, "treatment", "m.inc.d8", "m.n2",  colorsvolcano) 
e <- plot.volcano(m.inc.d8.prolong, "treatment", "prolong", "m.inc.d8",  colorsvolcano) 
f <- plot.volcano(prolong.extend, "treatment", "extend",  "prolong",colorsvolcano)

pituitary <- plot_grid(a + 
                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "pituitary \n -log10(p)"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e + theme(axis.title = element_blank(), axis.text = element_blank()),
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1))
pituitary

background <- png::readPNG("../figures/images/DoveParentsRNAseq_manipvolcanos.png")
icons <- ggdraw() +  draw_image(background, scale = 1)

plot_grid(icons, pituitary, nrow = 2, rel_heights = c(1,1))
```
